Analysis of Global Crises and Their Impact on Interational Well-Being

### 2024-11-15 Draft.Rmd
deleted clutter

```{r insight2, fig.height=4, fig.width=6, fig.cap="Boxplot of perception of corruption vs total COVID cases by continent"}

@@

```{r insight2, fig.cap=""}
#Social support & life ladder

data_combined <- inner_join(covid_2020, happiness, by = "country")

median_social_support <- median(data_combined$social_support, na.rm = TRUE)

ggplot(data_combined, aes(x = social_support, y = life_ladder, color = continent)) +
  geom_point(alpha = 0.7, position = "jitter") + 
  geom_smooth(method = "lm", color = "blue", se = TRUE) + 
  geom_vline(xintercept = median_social_support, color = "red", linetype = "dashed", linewidth = 1) + 
  labs(
    title = "Comparison of Social Support and Life Ladder by Continent",
    x = "Social Support",
    y = "Life Ladder (Happiness Score)",
    color = "Continent"
  ) +
  facet_wrap(~ continent) + 
  theme_minimal() +
  theme(panel.grid.major = element_line(color = "grey", linetype = "dashed", linewidth = 0.5))

# Social Support & Life Expectancy

ggplot(data_combined, aes(x = social_support, y = healthy_life_expectancy_at_birth, color = continent)) +
  geom_point(alpha = 0.7, position = "jitter") + 
  geom_smooth(method = "lm", color = "blue", se = TRUE) + 
  geom_vline(xintercept = median_social_support, color = "red", linetype = "dashed", linewidth = 1) + 
  labs(
    title = "Comparison of Social Support and Life Expectancy by Continent",
    x = "Social Support",
    y = "Healthy Life Expectancy at Birth",
    color = "Continent"
  ) +
  facet_wrap(~ continent) + 
  theme_minimal() +
  theme(panel.grid.major = element_line(color = "grey", linetype = "dashed", linewidth = 0.5))

data_combined <- inner_join(covid_2020, happiness, by = "country")

median_total_cases <- median(data_combined$total_cases, na.rm = TRUE)

ggplot(data_combined, aes(x = total_cases / 100, y = freedom_to_make_life_choices)) +
  geom_point(alpha = 0.5, position = "jitter") + 
  geom_smooth(method = "lm", color = "blue", se = TRUE) + 
  geom_vline(xintercept = median_total_cases, color = "red", linetype = "dashed", linewidth = 1) + 
  labs(
    title = "Comparison of COVID-19 Cases and Freedom",
    x = "Total COVID-19 Cases",
    y = "Freedom"
  ) +
  scale_x_log10() +  
  scale_y_log10() + 
  theme_minimal() +
  theme(panel.grid.major = element_line(color = "grey", linetype = "dashed", linewidth = 0.5))

#Life expectancy & Cyclone count

cyclone_counts <- cyclones_data |>
  group_by(country) |>
  summarise(cyclone_count = n())

combined_data <- cyclone_counts |>
  inner_join(happiness, by = "country") |>
  select(country, cyclone_count, healthy_life_expectancy_at_birth)

combined_data <- combined_data |>
  arrange(desc(cyclone_count)) |>
  slice(-1:-3)

ggplot(combined_data, aes(y = cyclone_count, x = healthy_life_expectancy_at_birth)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Relationship Between Cyclone Count and Life Expectancy (Outlier Removed)",
    x = "Cyclone Count",
    y = "Healthy Life Expectancy at Birth"
  ) +
  theme_minimal()

colnames(happiness)

# distance & cyclone count

# Load required libraries
library(tidyverse)

# Step 1: Aggregate cyclone counts by country
cyclone_counts <- cyclones_data |>
  group_by(country) |>
  summarise(cyclone_count = n())

# Step 2: Join with happiness data on 'country' to add distance from the equator
combined_data <- cyclone_counts |>
  inner_join(happiness, by = "country") |>
  select(country, cyclone_count, distance_to_equator_km)

# Step 3: Remove the top 3 highest cyclone counts
combined_data <- combined_data |>
  arrange(desc(cyclone_count)) |>
  slice(-1:-3)

# Step 4: Plot the relationship between distance to equator and cyclone count
ggplot(combined_data, aes(x = distance_to_equator_km, y = cyclone_count)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(
    title = "Comparison of Cyclone Count and Distance from Equator (Top 3 Outliers Removed)",
    x = "Distance from Equator (km)",
    y = "Cyclone Count"
  ) +
  theme_minimal()

ggplot(cyclones_data |> filter(!is.na(category)), aes(x = km_distance_from_equator, y = category, fill = category)) +
  geom_density_ridges(scale = 2, alpha = 0.7, color = "black", size = 0.3) +
  labs(
    x = "Distance from Equator (km)",
    y = "Cyclone Category",
    title = "Ridgeline Plot of Distance from Equator by Cyclone Category"
  ) +
  coord_cartesian(xlim = c(0, 6500)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

colnames(happiness)

ggplot(cyclones_data |> filter(!is.na(category)), aes(x = km_distance_from_equator, y = category, fill = category)) +
  geom_density_ridges(scale = 2, alpha = 0.7, color = "black") +
  annotate("rect", xmin = 1500, xmax = 2500, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.8) +
  labs(
    x = "Distance from Equator (km)",
    y = "Cyclone Category",
    title = "Ridgeline Plot of Distance from Equator by Cyclone Category"
  ) +
  coord_cartesian(xlim = c(0, 6500)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

happiness <- happiness|>
  left_join(
    rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") |>
      transmute(
        country = name_long,
        distance_to_equator_km = geosphere::distGeo(
          cbind(st_coordinates(st_centroid(geometry))[,1], st_coordinates(st_centroid(geometry))[,2]),
          cbind(st_coordinates(st_centroid(geometry))[,1], 0)
        ) / 1000  # Convert to km
      ),
    by = "country"
  )

happiness

>>>>

```{r insight2, fig.height=4, fig.width=6, fig.cap=""}

@@
